%%%%%%%%%%%%%%
% Robustness %
%%%%%%%%%%%%%%

\section{Robustness}

\subsection{Lag Structure and Consistency of Effects}

To test the robustness of the main findings, we estimate a series of fixed-effects models with lagged 
ICT investment variables from 1 to 8 years. The goal is to assess whether the impact of ICT 
investment on unemployment is stable over time and whether delayed effects differ across educational 
groups.

Across all lag structures, the core results hold. ICT investment remains positively and 
significantly associated with unemployment, particularly for low-skilled workers. The size and 
significance of coefficients are remarkably consistent. For example, the effect of a 3-year lagged 
ICT investment on unemployment is 4.846 (***), 2.935 (***), and 1.265 (***) for low-, medium-, and 
high-skilled workers, respectively (see Table~\ref{tab:models_3ylag}).

Interaction effects with welfare regimes are similarly robust. In all lag specifications, ICT 
investment has significantly weaker effects in post-socialist countries (e.g., -5.557*** in the 
3-year lag model for low-skilled) and significantly negative effects for medium- and high-skilled 
workers in Southern European regimes. The Nordic regime shows small but positive effects on 
high-skilled unemployment, potentially reflecting the rapid pace of technological diffusion.

The models include country and year fixed effects, reducing the risk of omitted variable bias. The 
reduction in observations for longer lags is due to the truncation of early years in the panel and 
is not systematically related to the dependent variable, preserving the validity of the estimates.

The consistent direction, magnitude, and statistical significance of the results across all eight lag 
structures confirm that the relationship between ICT investment and unemployment is not driven by 
model specification or short-term fluctuations. These findings support the temporal robustness of the 
digitalization–unemployment link.

Additional robustness checks—such as excluding the global financial crisis years or using alternative 
codings of welfare regimes—are discussed in the Appendix and confirm the stability of the results.

\subsection{Reverse Causality and Temporal Ordering}

One concern in assessing the relationship between ICT investment and unemployment is the possibility 
of \textit{reverse causality}: countries experiencing rising unemployment may respond with increased 
public or private ICT investments as part of structural reform strategies. If this is the case, 
observed associations could partially reflect policy reactions rather than causal effects of 
digitalization on employment.

To address this issue, we employed time-lagged models with lags from 1 to 8 years. These lagged 
specifications ensure that ICT investments precede labor market outcomes in time, thus mitigating 
simultaneity bias. The consistency of results across different lag lengths, especially the robustness 
of the 3-year lag model, supports the interpretation that the effect runs from ICT investment to 
unemployment, not the other way around.

Nevertheless, full elimination of endogeneity cannot be achieved in observational settings. Future 
research should consider instrumental variable approaches or quasi-experimental designs to strengthen 
causal inference.
